\documentclass[12pt]{letter}
\usepackage{fullpage,times}

%\pagestyle{empty}

\begin{document}

\begin{letter}{North Country This Week\\19 Depot St.\\Potsdam, NY}

\address{93 Elm St.\\Potsdam, NY 13676}

\signature{Jason Rohrer}


\opening{To the Editor:}

You know that Potsdam guy on Elm Street with the tall grass?  I'm that guy.

You may have read stories and seen pictures in the newspapers.  There was a ticket, there was a trial, and there was a verdict in our favor.  Now there is an appeal.

Most Village residents believe in a simple principle:  ``to each his or her own.''  However, a few people are upset.  Why are we doing this?  Just to make people angry?  Because we are lazy?  I hope that this letter will help you understand our position.

First of all, what are we trying to do on our property?  Instead of cultivating a lawn, we are trying to cultivate a natural wildflower meadow.  If you travel down any country road, you will see many beautiful meadows.  That kind of meadow is our eventual goal.

What have we done so far?  In 2004 and 2005, we physically removed much of the existing grass in the front yard by rototilling, hand raking, and mulch smothering.  We have planted coneflowers and sunflowers.  We have pulled thousands and thousands of ``weeds'' that we do not want, including Ragweed, which causes allergies, and the invasive Purple Loosestrife.  We have constructed a walking trail with signs identifying native wildflowers.  We have trimmed down the corner plants to give motorists a clear line of sight.

Why are we doing this?  Speaking for myself, I simply like the way a meadow looks.  Natural variety is beautiful to me.  My landscaping choice is part of who I am, and I cannot imagine myself having a lawn.  If you have a lawn and like it, that's fine with me---it's your property, and I support your choice.  In fact, if you like lawns, you're lucky, because you can cultivate your chosen landscape without risk of jail time.

I have now been ``in court'' for more than a year.  This is a criminal proceeding, not a civil lawsuit, and I face up to 15 days in jail.  I am representing myself and have already spent hundreds of hours doing legal research and work.  The first trial, on December 15, 2005, lasted three hours.  Six months later, we got a verdict of ``not guilty.''  Now, with the pending appeal, I must prepare for round two.  I'm certainly not going through this exhausting process just to defend my ``right to be lazy.''

Here's the bottom line:  should we, as a nation of free people, have the right to landscape our own properties however we choose, or should the government be able to force one style of landscaping on everyone?  You know what I believe.


\closing{Sincerely,}

\end{letter}

\end{document}

