\documentclass[12pt]{letter}
\usepackage{fullpage,times}

%\pagestyle{empty}

\begin{document}

\begin{letter}{Exodus Trust Archives\\1523 Franklin Street\\San Francisco, CA 94109}

\address{93 Elm St.\\Potsdam, NY 13676}

\signature{Jason Rohrer}


\opening{To the Archive maintainers:}

I am an artist, activist, writer, technologist, and parent.
When searching for educational materials that depict human sexuality in a realistic and honest fashion, I am dismayed to find that no such materials are readily available.

I have read quite a bit about the educational films produced by Ted McIlvenna and the National Sex Forum in the 1970s.
However, I have not been able to find these films, or video copies of them, archived in any library that I have searched.

In other words, these honest depictions of human sexuality have apparently been wiped from the face of the earth.
In a recent conversation that I had with Dr.\ McIlvenna, he chalked up the disappearance of such materials to the so-called Helms Amendment (which denies federal funding to organizations that produce sexually explicit materials).

However, Dr.\ McIlvenna also told me that many of his films from the Forum have been transfered to DVD as part of a collaboration between the Exodus Trust and the soon-to-open Erotic Heritage Museum.

I am interested in receiving samples of these forthcoming DVD materials.
I would be willing to write online reviews of any materials that you send me,  and I will also try to get the materials added to our local library's archives. 
Here is my address:
\begin{center}
\begin{tabular}{l}
Jason Rohrer\\
93 Elm St.\\
Potsdam, NY 13676 
\end{tabular}
\end{center}

I find your efforts, including the continued work of Dr. McIlvenna, to be a ray of hope.
The rest of our society seems to be marching, lock-step,  toward a new sexual dark age.
Godspeed in fighting the good fight, and good luck to you with the opening of the Heritage Museum.

\closing{Sincerely,}

\end{letter}

\end{document}

