\documentclass[12pt]{article}
\usepackage{fullpage,times}

%\pagestyle{empty}

\begin{document}

{\Huge

\begin{itemize}

\item Question:  can games be art?  I'm done asking that question.  How do we make art with games? [Slide: title]

\item I want you to know what I mean by ``art.''  Many people give up on defining it.  I'll propose a weak def.  [Slide: weak vs strong def] [Slide: weak def]

\item Comparing games and films

\item Wrong-headed comparisons with film [Slide: Ebert w/quote] [Slide: Ebert w/horns]

\item Study differences between games and films

\item Film toolbox.  Resonance is key.

\item Memento example.  What's the message?  [Slide:  Memento still]

\item Memento:  look for resonance

\item How are games different from film?  Gameplay is key.

\item What is gameplay?  Game mechanics?  Pac-Man example.  [Slide: pac-man]

\item Game without gameplay?  Beating the ``gameplay is key'' drum.

\item Gameplay rarely resonates.  Gameplay as forethought.  Same gameplay, different messages.  Genres.  [Slide:  game vs. film genres]

\item Imagine Memento if the film mechanics came first.  Backwards, but we do this with games, pick mechanic first.  [Slide:  Memento still]

\item Games try to express observations about human condition.  Cutscenes, linear stories.  MGS 1 and 2.  Different messages, exactly the same gameplay. [Slide:  Snake vs. Snake2] [Slide:  Raiden]

\item Ballroom dance scene vs. street rumble and camera motion

\item Dissonance on purpose?

\item Gameplay dissonance worse in games, because gameplay is heart.

\item Knocking BioShock.  The most inspired game environment ever.  Made me think.  It is a work of art. [Slide:  BioShock 1]

\item But: gameplay is stuck in FPS rut, doesn't resonate.  List mechanics (shooting stuff, key-and-lock puzzles).  [Slide: BioShock 2]

\item Why?  Because they set out to make a FPS first, came up with artistic expression later.  [Slide:  Ken Levine quote]

\item Challenge:  find resonant mechanics.

\item Rely on tools from other mediums less and less.

\item Films rely on text less and less (old days of film).  [Slide:  old film title card, 1915]
 
\item Core gameplay lacks human interest, so we resort to cut scenes.  The old days of game design?

\item  How far are we away from this?  A literal handful of people are starting to do this already.
 
\item Discuss three example game briefly.  Ignoring my own games, these are the only games that I know of that use gameplay as a vehicle for artistic expression. 
 
\item First:  The Marriage. [Slide:  The Marriage]

\item Braid [Slide: Braid Title]
 
\item Super Columbine Massacre RPG [Slide: SCMRPG]

\item Three example games, three routes through the creative process.
 
\item Where to go next?  How to craft resonant mechanics?
 
\item Learning the limitations of our medium.  Follow the strengths of mechanics.
 
\item Some topics might not be game-worthy, some topics can be dealt with better by games than other mediums

\item Example:  Passage and Memento Mori [Slide: memento mori painting] [Slide:  Passage]

\item What is memento mori really about?

\item Lesson:  games beat other mediums when dealing with processes and choices

\item Can't make a good game that's just about the loss of a loved one.

\item Gameplay that resonates can be a tool to weed out game-unworthy topics.

\item Artgames is a wide-open field with very few players.  Welcome.

\end{itemize}
}

\end{document}