\documentclass[12pt]{article}
\usepackage{fullpage,times,verbatim}

%\pagestyle{empty}


\begin{document}

\begin{center}
2014 vs. 2018:

The Shape of Financial Success Before and After the Indiepocalypse

Jason Rohrer

Independent Games Summit

Game Developers Conference 

March 18, 2019
\end{center}


{\Huge



\begin{itemize}

\item Talk is all about finanical results

\item Depending on lens, either impressive or mediocre.

\item Some of you just getting started in career

\item Don't want to give false impression

\item Comparing 17th game to 19th game [Slide:  17 v 19]

\item Big picture (lifetime actually in the bank money) [Slide: big picture]

\item First time ever calculated this total

\item First reaction:  Wow.

\item Hold on.  Been doing this 15 years.  Made 19 games.  [Slide: 15 years]

\item No money for 6 years [Slide]

\item No money until 14th game [Slide]

\item Repeat that:  I made 13 games that didn't make money [Slide]

\item Good money now?  Yes!  [Slide]

\item More than I could make in any other job.

\item But... [Slide: 3.5 years]

\item Per year [Slide:  ave per year]

\item Per game [Slide:  ave per game]

\item Could have made more over time with a job [Slide:  apple]

\item Long term project [Slide: long term]

\item Success takes time to build

\item Getting good at making games takes time

\item  Expect lots of failure

\item Don't give up.

\item Even after success, failure can come [Slide: game 18]

\item Alternate title [Slide: funny title]

\item End of prologue [slide]

\item Where did this indiepocalypse idea come from? [Slide: the word]

\item A string of publicly announced, high profile failures

\item From well-known devs.  [Slides:  Az, Wine, Scanner, Tacoma]

\item And 3.5 years into OHOL dev, this stuff made me nervous.

\item But somehow [Slide:  OHOL Steam]

\item And old game that's still making money [Slide:  Age and Players]

\item Concurrent players at an off time [Slide: monday]

\item Luck?  [Slide]

\item But looking through SteamCharts, I'm finding some other paid games with even more players.

\item Ever heard of this?  [Slide:  Staxel]

\item Or this?  [Slide: vineland]

\item Going higher, what about this? [Slide: airport]

\item Or this? [Slide: Fish]

\item Maybe you've heard of some of these games [Slide: 25 game]

\item What about these games: 50 game]

\item Or these? [Slide: 100 game]

\item Or these? [Slide: 200 game]

\item All of these have MORE concurrent players than OHOL

\item So far, I've put 375 games up on the screen, but that's only a fraction.

\item Actually 1050 [Slide:  1050]

\item So there are TONS of success stories out there

\item Maybe more than ever before.

\item We just don't hear about them.

\item Still, a lot has changed since I launched my last game [Slide: change table]

\item Steam crowding [Slide]

\item No one covered OHOL [Slide]

\item No one covered OHOL [Slide]

\item Dismal units [Slide]

\item Dismal dollars [Slide]

\item All these numbers are an order of magnitude better for TCD

\item Really bad for OHOL.

\item First month curve TCD [Slide]

\item We've all seen this curve

\item Expected curve for game sales

\item Launch day peak with expoential decay

\item Scary thing:  Launch peak of OHOL so much lower [Slide]

\item If expected curve kicks in, we're doomed [Slide]

\item Shocked when day 2 was higher [Slide]

\item And day 3 even higher [Slide]

\item Finally reached peak on day 5, which was actually a Monday [Slide]

\item And we might expect decay curve to kick in [Slide]

\item But the actual curve looks substantially different [Slide]

\item How to account for this?

\item Look at some other launch week differences [Slide]

\item Almost no one made YT vids for TCD [Slide]

\item Almost nobody watched them [Slide]

\item Overall compare [Slide]

\item Overall, huge YT views difference [Slide]

\item Notice that these numbers are an order of magnitude better for OHOL.

\item So what does this mean?

\item This is a press curve [slide]

\item Press is dead [Slide]

\item Tons of press for Wine [Slide]

\item Even more for Tacoma [Slide]

\item But nobody read any of it

\item I got no press, but it didn't matter [Slide]

\item Here's a game with absolutely no press [Slide]

\item Normal hype curve continues like this in month 2 [Slide]

\item OHOL started month 2 higher, so that's good [Slide]

\item The expected decay curve would look like this [Slide]

\item What actually happened is shocking [Slide]

\item Press, even if people read it, can't do this

\item This is a sign of word of mouth.

\item YT videos are one kind of word of mouth.

\item Now look at these games with disappointing sales [Slide:  wine YT]

\item Tacoma did a bit better, but not by much [Slide]

\item Notice the short running times on these videos

\item These are review videos

\item But look at the top video for OHOL [Slide]

\item 24 minutes long, and it's just the guy playing the game [Slide]

\item What about an even more popular game [Slide]

\item Again, another video with someone just playing the game

\item What about other kinds of word of mouth? [Slide:  month 2]

\item What's that spike there?  Xmas day [Slide]

\item Word of mouth only works if people still playing your game in 4 months.

\item Sample from hours listed on front-page Steam reviews [Slide: Wine]

\item These are short games [Slide: tacoma]

\item This is long game [Slide: OHOL]

\item So is this one [Slide: Fish]

\item Some people may have gone overboard here [Slide: max hours OHOL]

\item While a lone fanatic may help with word of mouth, span is more important.

\item People play OHOL over an extended period of time [Slide: OHOL]

\item Some have been playing it for a year [Slide: OHOL max]

\item Good portion over 4 months [Slide: OHOL 10%]

\item Kinds of games have changed [Slide: from/to]

\item Way players think about them has changed [Slide:  Do you play?]

\item But calling this new is kinda strange [Slide:  chess]

\item Advice needs to change [Slide: actually play]

\item Look at your own play history [Slide: steam library]

\item Really liked Edith Finch [Slide: finch]

\item But it didn't become part of my life [Slide: rust]

\item More chance of recommending it [Slide:  finch rec]

\item People kept hearing me talk about Rust [Slide:  rust rec]

\item My summary of this difference [Slide: consumable vs generators]

\item Look at all the indie games that have been huge hits.

\item Factorio, Subnautica, Stardew Valley, Prison Architect.

\item And this fits in with YT videos [Slide: twisted]

\item How else would 159 videos about one game be possible?

\item And it's no coincidence, high profile games consumable [Slide]

\item While the no-name hit games are all unique situation generators [Slide]

\item So the Indieapocalypse isn't real. [Slide: strikethrough]

\item Consumable-gamepocalypse [Slide]

\item Take-away:  MAKE UNIQUE SITUATION GENERATORS [Slide]

\item Aren't those the most interesting kinds of games, anyway?

\end{itemize}

}

\end{document}
