\documentclass[12pt]{article}
\usepackage{fullpage,times,amssymb}

\renewcommand{\labelitemi}{$\square$}
\renewcommand{\labelitemii}{$\square$}


\newcommand{\M}{Monday}
\newcommand{\Tu}{Tuesday}
\newcommand{\W}{Wednesday}
\newcommand{\Th}{Thursday}
\newcommand{\F}{Friday}


\newcommand{\dailyChecklist}[1]{ 

\subsection{#1}
\begin{itemize}
\large

\item Draw a detailed picture of something from the real world.
\item Typing lesson.
\item Draw and develop one new comic character.
\item Compose and practice a new piece of piano music.
\item Read next chapter of {\it Understanding Comics}.
\item Apply ideas from {\it UC} chapter to draw a new, simple 4-panel stick figure comic.
\item Japanese lesson.
\item Next step in short film work.
\end{itemize}

}

\newenvironment{planDay}[1]{\subsection{#1}\begin{itemize} \large }{\end{itemize}}


\begin{document}


\section{Daily Plans}
\dailyChecklist{\M}
\dailyChecklist{\Tu}
\dailyChecklist{\W}
\dailyChecklist{\Th}
\dailyChecklist{\F}



\newpage

\section{Typing Lessons}

\begin{planDay}{\M}
\item a s d f
\item j k l ;
\end{planDay}

\begin{planDay}{\Tu}
\item e i
\item g h
\end{planDay}


\begin{planDay}{\W}
\item r u 
\item t y
\end{planDay}

\begin{planDay}{\Th}
\item q w o p
\item v b m n
\end{planDay}

\begin{planDay}{\F}
\item z x c 
\item . ,
\end{planDay}




\newpage

\section{Japanese Lessons}

\begin{planDay}{\M}
\item Unit 1, Lesson 1 (all)
\end{planDay}

\begin{planDay}{\Tu}
\item Unit 1, Lesson 2 (parts 1, 2)
\end{planDay}

\begin{planDay}{\W}
\item Unit 1, Lesson 2 (parts 3, 4, 5)
\end{planDay}

\begin{planDay}{\Th}
\item Unit 1, Lesson 3 (parts 1, 2, 3)
\end{planDay}

\begin{planDay}{\F}
\item Unit 1, Lesson 3 (parts 4, 5, 6, 7, 8, 9)
\end{planDay}



\newpage

\section{Short Film Plan}

\begin{planDay}{\M}
\item Read screenplays.
\item Pick a screenplay to produce.
\end{planDay}

\begin{planDay}{\Tu}
\item Contact actor friends.
\item Deliver script to actor.
\item Rehearse your part at home.
\item Compose shot list.
\item Storyboard sketches for needed shots.
\end{planDay}

\begin{planDay}{\W}
\item Meet with actor friend.
\item Rehearse lines and blocking.
\item Plan shots.
\item Dry run with tripod.
\end{planDay}

\begin{planDay}{\Th}
\item Principal Photography.
\item Follow shot list.
\end{planDay}

\begin{planDay}{\F}
\item Import footage into computer.
\item Rough edit.
\end{planDay}



\end{document}

