\documentclass[12pt]{article}
\usepackage{fullpage}

\textwidth 5.7in
\marginparsep .25in

% re-adjust margin note width to work with text width
\setlength{\marginparwidth}{8.5in}
\addtolength{\marginparwidth}{-1.0in}
\addtolength{\marginparwidth}{-\oddsidemargin}
\addtolength{\marginparwidth}{-\textwidth}
\addtolength{\marginparwidth}{-2.0\marginparsep}
% a little extra is necessary
\addtolength{\marginparwidth}{-1\marginparsep}


\newcommand{\sidenote}[1]{\marginpar{\begin{flushleft}{\sc \hspace{0pt}#1}\end{flushleft}}}

\begin{document}


\title{Purchase/Sale Agreement}

\maketitle

\begin{enumerate}

\item This \sidenote{dates and parties} agreement is made on \today \  between {\bf Jason Rohrer and Lauren Serafin} (the {\it Sellers}) and  {\bf Brian Huss and Kristin Andrews} (the {\it Buyers}).
\begin{center}
\begin{tabular}{|l|l|}
\hline
Sellers' Contact Information: & Buyers' Contact Information:\\
\hline
93 Elm St.                    & 9 Leroy St.\\
Potsdam, NY 13676             & Potsdam, NY 13676\\
315-265-0585                  & 315-262-6078\\
                              & \\
                              & 463 Brock Ave.\\
                              & Toronto, ON M6H 3N8\\
                              & 416-535-5793\\
                     
\hline
\end{tabular}
\end{center}

\item The \sidenote{property description} Sellers agree to sell and convey to the Buyers, and the Buyers agree to purchase from the Sellers, the real estate, together with all buildings and improvements, located at 93 Elm St. in the Potsdam Village, St. Lawrence County, New York State (the {\it Property}).  The Property includes a house with a one-car garage, a storage shed, and a lot with 123 feet of street frontage.  The Tax Map number is {\bf 64.060-4-9}. 

\item The \label{ref_payment} \sidenote{price} purchase price is sixty-five thousand dollars ({\bf \$65,000}), which will be paid by the Buyers to the Sellers as follows:

\begin{enumerate}
\item an initial cash deposit of ({\bf \$2000}) and 
\item the remaining balance at closing ({\bf \$63,000}).
\end{enumerate}

All money deposited prior to closing will be paid to the Sellers, who shall retain the deposits until the completion or termination of this agreement.

%This Agreement represents a {\bf cash offer} and is not contingent upon the Buyers obtaining a mortgage.  For details, see part \ref{ref_contingencies}.


\item This \sidenote{time limit} offer is good until {\bf April 15, 2009 at noon}.  Should the offer not be accepted, all deposit money will be returned to the Buyers.

\item Closing \sidenote{closing} will be made at the office of the Sellers' attorney, or at another location chosen by the Sellers, on or before {\bf July 20, 2009}, or as soon as the abstract can be brought up to date, but in no event later than sixty (60) days after the scheduled closing date unless this Agreement is extended in writing by all parties.

\item Possession \sidenote{possession} of the Property will be delivered at closing.

\item The \sidenote{title and abstract} Sellers will convey the Property free and clear of all liens and encumbrances, except for existing building restrictions, zoning restrictions, ordinances, easements, restrictions of use, and any encumbrances that a physical inspection of the Property would reveal.  The title will be good and marketable.

At least ten (10) days before the closing date, the Sellers will deliver to the Buyers a forty-year abstract of title certified down to date and ten year tax search.  If the title to the Property is found to be defective in the opinion of the Buyers, the Sellers will have thirty (30) days after receipt of written notice within which to fix the defects.  At the time of the closing, the title will be conveyed by warranty deed with lien covenant.


\item If \sidenote{survey} any surveys or deed plots are deemed necessary because the existing description of the Property is unascertainable, they will be obtained and paid for the the Sellers.

\item All \sidenote{adjustments} taxes, water and sewer assessments, rents and fuels remaining on the Property are to be pro-rated as of closing. State, County and Town Taxes will be adjusted and apportioned on a calendar year beginning January 1 and ending December 31. School Taxes will be adjusted and apportioned for the fiscal year beginning July 1 and ending the following June 30, and Village Taxes will be adjusted and apportioned for the fiscal year beginning June 1 and ending the last day of May following or as otherwise provided by law.

\item The \sidenote{inclusions} following items, unless specifically excluded from this Agreement, will become the property of the Buyers and are included in the purchase price:  

\begin{enumerate}
\item all plumbing, heating, and lighting fixtures and systems;
\item all built-ins, laundry tubs, disposals, TV antennas, TV rotors, door knockers, fire alarms, smoke alarms, burglar alarms, mailboxes, weather vanes, and other permanent fixtures;
\item all screens, storm windows and doors, venetian blinds, shades, awnings, valances, curtain rods, drapery rods, and traverse rods; and
\item all trees, shrubs, and plantings now on the Property.
\end{enumerate}

The following personal property is also included in the purchase price:
\begin{enumerate}
\item reverse osmosis filter for drinking water,
\item stove,
\item refrigerator,
\item dishwasher, and
\item clothes washer.
\end{enumerate}

Prior \sidenote{exclusions} to the July 20, 2009 possession date, all other personal property will be removed from the Property, including items in the attic, garage, and storage areas, at the expense of the Sellers.


\item The \sidenote{damage or loss} parties agree that the provisions of the uniform Vendor and Purchaser Risk Act (New York General Obligations Law, 5-1311) will govern the risk or loss in the event that, prior to closing, the Property or any part of it is:
\begin{enumerate}
\item destroyed by fire or another cause or
\item taken for public use.
\end{enumerate}


\item In \sidenote{default} the event that the Sellers are unable to give good and marketable title, the Buyers will have the option of taking such title as the Sellers can give without reduction of the purchase price, or of being repaid all money paid by the Buyers and held by the Sellers as part of the purchase price, in which case the Buyers will be released from all liability, and this Agreement will become null and void.  

Should the Buyers fail to make additional payments as specified in part \ref{ref_payment}, or violate or fail to fulfill and perform any of the terms or conditions of this Agreement, all deposits and other sums paid by the Buyers on account of the purchase price, whether required by this Agreement or not, may be retained by the Sellers, either on account of the purchase price or as liquidated damages for such a breach, as the Sellers may elect, and in the latter event, the Sellers will be released from all liability and this Agreement will become null and void.  The Sellers' waiver of one breach of any covenant on this contract will not be construed as a waiver of any subsequent breach.

\item The \sidenote{representations} Buyers have inspected the Property and have agreed to purchase it as a result of such inspection and not because of or in reliance upon any representation made by any agent of the Sellers.  However, the Sellers affirm that the Sellers have disclosed any and all known latent defects in the structure or operation of the Property.  The Buyers have agreed to purchase the Property in its present condition (``as is'') unless otherwise specified in this Agreement.  All appliances will be sold ``as is.''  The Sellers agree to maintain the Property's grounds, improvements, and buildings in the same condition as they are at the time that this Agreement is signed, reasonable wear and tear excepted.  The Buyers affirm that they are aware of all zoning restrictions that would be applicable to the Property.

\item This \sidenote{whole agreement} Agreement contains the whole Agreement between the Sellers and the Buyers, and there are no other terms, obligations, covenants, representations, statements, or conditions, oral or otherwise of any kind whatsoever concerning this sale.  Any changes or conditions to this Agreement must be made in writing and executed by the parties described in this Agreement.

\item This \sidenote{recording} Agreement will not be recorded in the County Clerk's Office or in any other office of public record.

\item This \sidenote{assignment} Agreement will be binding upon the respective heirs, executors, administrators, distributees, successors, and assigns of the parties described in this Agreement.

\item This \label{ref_contingencies} \sidenote{contingencies} Agreement not subject to any contingency clauses for the following reasons:
\begin{enumerate}
\item This Agreement is not contingent upon the Buyers obtaining a mortgage.  The Buyers affirm that they have an alternative funding source sufficient to fulfill the payment terms as specified in part \ref{ref_payment}.
\item  The Buyers have already completed a house inspection and have received a satisfactory inspection report.
\end{enumerate}


\item This \sidenote{legally binding} is a legally binding document when signed.  It imposes specific legal obligations and gives specific legal rights to both the Buyers and the Sellers.  If you do not understand this agreement or any portion of it, you should consult an attorney before signing.

\end{enumerate}


\vspace{.5in}

\begin{center}
\begin{tabular}{p{.5in}p{2in}p{.25in}p{.5in}p{2in}}
\hrulefill&\hrulefill&&\hrulefill&\hrulefill\\
{\sc date}&{\sc buyer}&&{\sc date}&{\sc buyer}\vspace{.5in}\\
\hrulefill&\hrulefill&&\hrulefill&\hrulefill\\
{\sc date}&{\sc seller}&&{\sc date}&{\sc seller}
\end{tabular}
\end{center}


\end{document}

