\documentclass[12pt]{article}
\usepackage{fullpage,times}

%\pagestyle{empty}

\begin{document}

{\Huge

\begin{itemize}

\item How do we make art with games? [Slide: title]

\item Comparing games and films

\item Wrong-headed comparisons with film [Slide: Ebert w/quote] [Slide: Ebert w/horns]

\item Study differences between games and films

\item Film toolbox.  Resonance is key.

\item Memento example.  What's the message?  [Slide:  memento poster]

\item Memento:  look for resonance

\item How are games different from film?  Gameplay and game mechanics are key.

\item Game without gameplay?  Beating the ``gameplay is key'' drum.

\item Gameplay rarely resonnates.  Gameplay as forethought.  Same gameplay, different messages.  Genres.  [Slide:  game vs. film genres]

\item Imagine Memento if the film mechanics came first.  Backwards, but we do this with games, pick mechanic first.

\item Gameplay dissonance worse in games, because gameplay is heart.

\item Challenge:  find resonant mechanics.

\item Rely on tools from other mediums less and less.

\item Films rely on text less and less (old days of film).  [Slide:  old film title card, 1915]
 
\item Core gameplay lacks human interest.  The old days of game design?
 
\item My artgame list (Braid, The Marriage, SOHMB, SCMRPG)
 
\item Where to go next?  How to craft resonant mechanics?
 
\item Learning the limitations of our medium.  Follow the strengths of mechanics.
 
\item Some topics might not be game-worthy, some topics can be dealt with better by games than other mediums

\item Example:  Passage and Memento Mori [Slide: Passage + memento mori painting]

\item What is memento mori really about?

\item Lesson:  games beat other mediums when dealing with processes and choices

\item Can't make a good game that's just about the loss of a loved one.

\item Gameplay as metaphor as a tool to weed out game-unworthy topics.

\end{itemize}
}

\end{document}