\documentclass[12pt]{article}
\usepackage{fullpage, times, epsfig}



\pagestyle{empty}


\setlength{\textheight}{10in} \setlength{\textwidth}{7.5in}
\setlength{\oddsidemargin}{-0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}


% macro so we can repeat instructions twice on same sheet
\newcommand\instructions{
\begin{center}
INSTRUCTIONS
\end{center}

\begin{enumerate}
\item Only registered voters in the Town of Potsdam can sign or help gather signatures.
\item Petitions must be filled out {\bf in ink}.
\item You must personally witness every signature on your sheet.
\item You should fill in the Date, Print Name, and Residence fields for each signer to ensure that the printing is clear.  The signer just needs to sign---you can fill in the rest.
\item Signature dates on a given sheet must appear in order.
\item Wait to sign and date the witness statement until after you are finished with a sheet---no signature date should be later than the date of your witness signature date.
\item {\bf Do not} sign a signature line on your own sheet.  You can't witness for yourself. 
\item To correct a mistake in the Residence field, cross out the mistake with a single line and put your initials at each end of the line (your initials, not the signer's initials).
\item Leave the Sheet Number fields blank---they will be filled in when the petition sheets are bound together.
\end{enumerate}

You must return your completed petition sheets to me by {\bf Monday, February 4}:
\begin{center}
\begin{tabular}{c|c}

\begin{tabular}{l}
Jason Rohrer\\
93 Elm St.\\
Potsdam, NY 13676
\end{tabular}&

\begin{tabular}{l}
Questions?  Contact me at:\\
265-0585\\
jasonrohrer@fastmail.fm
\end{tabular}\\

\end{tabular}

\end{center}
}



\begin{document}

\instructions

\vfill

\instructions

\end{document}
