\documentclass[12pt]{article}
\usepackage{fullpage, times, epsfig}



\pagestyle{empty}


\setlength{\textheight}{10in} \setlength{\textwidth}{7.5in}
\setlength{\oddsidemargin}{-0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}


\begin{document}

\begin{center}
PETITION FOR TOWN REFERENDUM
\end{center}

I, the undersigned, do hereby state that I am a  registered  voter  of
 the  {\bf Town  of Potsdam} and that  my present  place of residence is truly
  stated opposite my signature.  I do hereby object to the bond resolution adopted by the Town Board on January 8, 2008 authorizing the issuance of bonds in amount not to exceed three million two hundred thousand dollars (\$3,200,000) for the construction, equipping, and furnishing of a new town hall and office building for the Township.  I request that this bond resolution be submitted to the qualified electors of the Town for their approval or disapproval, pursuant to Article 7, Section 91 of the New York State Town Law.

In witness whereof, I have signed this petition on the  day  and  year
  stated before my signature.

\vspace{0.25in}

{\Large

\newcounter{sigNumber}
\setcounter{sigNumber}{1}

\newcommand\sigLine{\thesigNumber. \addtocounter{sigNumber}{1} & & & &\\
\hline}

\begin{center}
\noindent\begin{tabular}{|r|p{0.08\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|}
\hline
&Date&Print Name&Residence&Signature\\
\hline
\hline
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\sigLine
\end{tabular}
\end{center}

}

\vspace{0.25in}

\begin{center}
STATEMENT OF WITNESS
\end{center}

    I,  \underline{\hspace{4in}}  state  that I am a duly qualified voter and
  now reside at  \underline{\hspace{4in}} in the {\bf Town of  Potsdam}  in  the  State  of  New  York  in  the  County  of
  St. Lawrence.  Each of the voters whose  names  are  subscribed  to
  this   petition  sheet, which contains \underline{\hspace{1in}} signatures, subscribed his or her name in my presence.
    
I understand that this statement will be accepted for all purposes  as
  the  equivalent  of  an  affidavit  and, if it contains a material false
  statement, shall subject me to the same penalties as if I had been  duly
  sworn.

\vspace{0.125in}

\begin{center}
\noindent\begin{tabular}{ccc}
\underline{\hspace{0.30\textwidth}} & \hspace{0.30\textwidth} & \underline{\hspace{0.30\textwidth}}\\ 
Date & & Signature of Witness
\end{tabular}

\vspace{0.25in}

Sheet Number \underline{\hspace{1in}}
\end{center}

\end{document}