\documentclass[12pt]{article}
\usepackage{fullpage,times,}

\pagestyle{empty}


\begin{document}

\begin{center}
{\Huge Photo Legend}
\end{center}

\noindent The following properties were photographed:
\renewcommand{\labelenumi}{(\Alph{enumi})}
\begin{enumerate}
\item the property between 101 and 103 Elm Street;
\item the property just beyond 102 Elm Street, up to the corner of Elm and Wellings Drive;
\item the property on the corner of Ridgewood Lane and Fairlawn Drive (across from the house at 10 Ridgewood and next to the house at 4 Ridgewood); and
\item the property just south of the house at 14 Fairlawn Drive.
\end{enumerate}
The photographs were numbered by the camera with the form {\bf 100-2774}.
For the sake of brevity, the numbers shown below omit the {\bf 100-} prefix, since it is common to the entire photograph set.



\paragraph{2744} property (A) from the west.
\paragraph{2745} a 15-inch ruler on the road in front of (A).
\paragraph{2746} detail of the ruler.
\paragraph{2747} the ruler held in the grass on property (A).
\paragraph{2748} a tape measure between the road and the grass on (A).
\paragraph{2749} detail of where the tape measure meets the tall grass.
\paragraph{2750} second detail of the tape measure.

\paragraph{2751} property (B) from the southwest.
\paragraph{2752} the rest of property (B) from the southwest.  Note that tall grass is present between the ditch and the street.
\paragraph{2754} property (B) from the south.
\paragraph{2756} Lauren Serafin (who is 5'10'' tall) standing in the grass on property (B) .
\paragraph{2757} property (D) from the west.
\paragraph{2758} Jason Rohrer (who is 6'8'' tall) and Mez (who was 36'' tall) standing in the grass on property property (D).
\paragraph{2759} property (C) from the east.  Note that the grass hangs over the road.
\paragraph{2760} property (D) from the southwest.  Note the close proximity of the tall grass to the neighboring house at 14 Fairlawn Drive.
\paragraph{2761} property (C) from the south.  Note the close proximity of the tall grass to the neighboring house at 4 Ridgewood Lane.
\paragraph{2762} the property at 4 Ridgewood Lane, which has a traditional landscape.  Note the shrubs which are trimmed into spiral formations.
\paragraph{2764} property (B) from the southeast.
\paragraph{2765} property (B) from the south.  Note the large house that is visible beyond the tall grass (the house shares the driveway at 102 Elm Street).
\paragraph{2766} a gravel path through the tall grass on property (B).
\paragraph{2767} property (B) from the southwest.  Note again the gravel path.
\paragraph{2768} property (B) in relation to 102 Elm Street.


\end{document}
