\documentclass[12pt]{article}
\usepackage{fullpage,times}


\begin{document}

\begin{center}
{\Huge Questions for Expert Witness Richard Grover}
\end{center}



\section{Foundation}

\subsection{Personal}

\begin{enumerate}
\item What is your name and current place of residence?
\end{enumerate}

\subsection{Education}
\begin{enumerate}
\item What is your educational history?
\end{enumerate}

\subsection{Professional experience}
\begin{enumerate}
\item What professional experience do you have in the area of Community Planning?
\item What professional experience do you have dealing with environmental issues?
\item Have you done any professional research concerning environmental issues?
\item What professional experience do you have in the area of landscape architecture?

\item How much of your professional experience as a landscape architect has dealt with natural landscaping?

\item Do you have any teaching experience?

\item What are your current titles and occupations?
\end{enumerate}


\subsection{Expert witness experience}
\begin{enumerate}
\item Have you ever served as an expert witness in a court of law?
\item Can you give a brief overview of the types of cases for which you served as an expert witness?
\end{enumerate}



\section{Expert Testimony}

\subsection{Environment}
\begin{enumerate}
\item What are some of the landscaping practices that are typical of traditional landscaping in our region?
\item What effect do these traditional landscaping practices have on the environment?
\item Are native plant species usually present in a traditional landscape?
\item What is your opinion on the importance of native plant species to the environment?
\end{enumerate}

\subsection{Expression}
\begin{enumerate}
\item Is landscaping an art form?
\item Is landscaping a form of personal expression?
\item As a landscape architect, do you attempt to communicate with the landscapes that you design?
\item What kind of message is delivered by your landscape design?
\end{enumerate}

\subsection{Natural landscaping}
\begin{enumerate}
\item In your professional work as a landscaper, have you ever helped to cultivate a meadow?
\item Can you describe the outcome of this process?
\item Did grass make up part of the meadow ecosystem?
\item Does grass play a significant role in the natural meadow habitats of our region?
\item Have you been paid for your work in installing meadows?
\end{enumerate}

\subsection{The 10-inch rule}
\begin{enumerate}
\item Did the meadow grass you described earlier exceed 10 inches in height?
\item Would regular mowing be necessary to keep meadow grasses from exceeding 10 inches in height?
\item What effect would regular mowing have on a natural meadow ecosystem?
\end{enumerate}

\subsection{Aesthetic}
\begin{enumerate}
\item Have you, in your professional work as a landscape architect, encountered people who find tall grasses to be aesthetically pleasing?
\end{enumerate}


\subsection{Hazards}
\begin{enumerate}
\item Are any hazards created by the cultivation of tall grasses?
\item Would you describe tall grass in our region as a serious fire hazard for neighboring properties and structures?
\item Have you ever encountered a grass fire in our region that damaged a nearby structure?
\end{enumerate}


\subsection{Changing ideas}
\begin{enumerate}
\item In your work as a community planner, have you observed cultural or political changes in terms of how citizens and governments perceive natural landscaping?
\item Are you aware of any government documents that encourage the cultivation of natural landscapes or native plants?
\end{enumerate}


\end{document}